---
title: "Family Poverty w/ Child Under 18 as an Outcome Variable | Clustering"
author: "Connor Flynn"
date: "9/21/2023"
output: html_document
---

Load in Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(plotly)
library(scales)
library(reactable)
library(tigris)
library(patchwork)
library(shiny)
library(leaflet)
library(tibble)  # For data manipulation
library(tidyr) 
library(sf)
library(RColorBrewer)
library(ranger)
library(missRanger)
library(randomForest)
library(here)
library(cluster)
library(factoextra)

options(scipen=99)
```


The American Community Survey (ACS) is conducted by the U.S. Census Bureau and collects detailed demographic, social, economic, and housing information from a sample of households in the United States. The "profile" dataset contains a wide range of demographic and socioeconomic variables, such as:

- Population Characteristics: Variables related to age, gender, race, ethnicity, marital status, and more.

- Housing Characteristics: Variables related to housing units, such as the number of rooms, tenure (renter-occupied or owner-occupied), and housing value.

- Economic Characteristics: Variables related to income, employment status, occupation, industry, and educational attainment.

- Social Characteristics: Variables related to disability status, veteran status, and more.

- Geographic Information: Variables related to geographic location, including state, county, tract, and block group.


Change to ACS 
```{r}
v16 <- load_variables(2021, "acs1/profile", cache = TRUE)
```


This pulls data specifically for the variable DP03_0120P (Percent!!PERCENTAGE OF FAMILIES AND PEOPLE WHOSE INCOME IN THE PAST 12 MONTHS IS BELOW THE POVERTY LEVEL!!All families!!With related children of the householder under 18 years)


```{r}
pct_families_income_under_poverty_level <- get_acs(
  geography = "tract", 
  variables = "DP03_0120P",
  state = "HI", 
  year = 2021,
  geometry = TRUE
)|>
  dplyr::filter(GEOID != "15003981200")
```

Plot Family Poverty by Census Tract

```{r}
ggplot(pct_families_income_under_poverty_level, aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "Family Poverty %")

```

Pull data on family overty over time in Hawaii


```{r}
years <- 2015:2021
names(years) <- years

pct_families_income_under_poverty_level_time_series <- map_dfr(years, ~{
  get_acs(
    geography = "tract",
    geometry = "true",
    state = "HI",
    variables = "DP03_0120P",
    year = .x
  )
}, .id = "year")

```
Drop census tract that contains northwestern Hawaiian Islands

```{r}
pct_families_income_under_poverty_level_time_series <- pct_families_income_under_poverty_level_time_series %>% 
  dplyr::filter(GEOID != "15003981200")
```



Plot Family Poverty over time

```{r}
ggplot(pct_families_income_under_poverty_level_time_series, aes(fill = estimate)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "Family Poverty %") +
  facet_wrap(~year)
```
Plot Distribution of Family Poverty over time

```{r}
ggplot(pct_families_income_under_poverty_level_time_series, aes(x = estimate)) +
  geom_histogram() +
  theme_minimal() +
  facet_grid(~year)
```
```{r}
pct_families_income_under_poverty_level_time_series <- pct_families_income_under_poverty_level_time_series %>%
  mutate(county = sub(".*?,\\s*(.*County, .*Hawaii)$", "\\1", NAME))
```


Add county info as color 

```{r}
ggplot(pct_families_income_under_poverty_level_time_series, aes(x = estimate, fill = county)) +
  geom_histogram(color = "black", bins = 50) +
  theme_minimal() +
  facet_grid(~year) +
  scale_fill_brewer(palette = "Set3")
```


Random Forest


Pull every variable from the ACS5 Profile Dataset for Hawaii by Census Tract (677)

```{r}

v16_percentages <- v16 %>%
  filter(grepl("^Percent", label))


all_variables <- as.list(v16_percentages$name)



# Use the list of loaded variables to fetch data for all census tracts in Hawaii
profile_percentage_variables_HI_tracts <- get_acs(geography = "tract", 
                         variables = all_variables,
                         state = "HI", 
                         year = 2021)
```


Drop Margin of Error Column and pivot wider so each variable becomes a column of the dataframe

```{r}
profile_percentage_variables_HI_tracts <- profile_percentage_variables_HI_tracts %>%
  select(-moe) %>% 
  pivot_wider(names_from = variable, 
              values_from = estimate)
```


Drop Columns with numeric values exceeding 100 (need to think through)

```{r}

# Define the columns to be checked for values exceeding 101 or NA values
columns_to_check <- setdiff(names(profile_percentage_variables_HI_tracts), "NAME")

# Loop through columns (excluding "NAME") and check for values exceeding 101 or NA values
columns_to_keep <- sapply(profile_percentage_variables_HI_tracts[columns_to_check], function(col) all(is.na(col) | col <= 101))

# Combine "NAME" column with columns meeting the condition
profile_percentage_variables_HI_tracts_true <- profile_percentage_variables_HI_tracts[, c("NAME", columns_to_check[columns_to_keep])]

```

Drop columns missing 30% or more of the data

```{r}
threshold <- 0.7

# Create the new dataframe by filtering columns
profile_percentage_variables_HI_tracts_true_less_na <- profile_percentage_variables_HI_tracts_true %>%
  select(NAME, where(~mean(!is.na(.)) >= threshold))
```

Save dataframe as hi_profile_data_cleaned

```{r}
hi_profile_data_cleaned <- profile_percentage_variables_HI_tracts_true_less_na
```

Impute missing data with MissRanger (takes some time)

```{r}
#hi_profile_data_cleaned_imputed <- missRanger(hi_profile_data_cleaned)
```

Save as csv in data folder

```{r}
# write_csv(hi_profile_data_cleaned_imputed, here("data/hi_profile_data_cleaned_imputed.csv"))
```

```{r}
hi_profile_data_cleaned_imputed <- read_csv(here("data/hi_profile_data_cleaned_imputed.csv"))
```
```{r}
ggplot(hi_profile_data_cleaned_imputed, aes(x = DP03_0120P)) +
  geom_histogram()
```
Drop poverty and Income Columns because they are too "obvious" of a predictor for poverty

```{r}
name_values <- c(
    "DP03_0121P",
    "DP03_0122P",
    "DP03_0123P",
    "DP03_0124P",
    "DP03_0125P",
    "DP03_0126P",
    "DP03_0127P",
    "DP03_0128P",
    "DP03_0129P",
    "DP03_0130P",
    "DP03_0131P",
    "DP03_0132P",
    "DP03_0133P",
    "DP03_0134P",
    "DP03_0135P",
    "DP03_0136P",
    "DP03_0137P",
    "DP03_0119P",
    "DP03_0094P",
    "DP03_0093P",
    "DP03_0092P",
    "DP03_0091P",
    "DP03_0090P",
    "DP03_0089P",
    "DP03_0088P",
    "DP03_0087P",
    "DP03_0086P",
    "DP03_0085P",
    "DP03_0084P",
    "DP03_0083P",
    "DP03_0082P",
    "DP03_0081P",
    "DP03_0080P",
    "DP03_0079P",
    "DP03_0078P",
    "DP03_0077P",
    "DP03_0076P",
    "DP03_0075P",
    "DP03_0074P",
    "DP03_0073P",
    "DP03_0072P",
    "DP03_0071P",
    "DP03_0070P",
    "DP03_0069P",
    "DP03_0068P",
    "DP03_0067P",
    "DP03_0066P",
    "DP03_0065P",
    "DP03_0064P",
    "DP03_0063P",
    "DP03_0062P",
    "DP03_0061P",
    "DP03_0060P",
    "DP03_0059P",
    "DP03_0058P",
    "DP03_0057P",
    "DP03_0056P",
    "DP03_0055P",
    "DP03_0054P",
    "DP03_0053P",
    "DP03_0052P",
    "DP03_0051P"
)

hi_profile_data_cleaned_imputed_dropped_features <- hi_profile_data_cleaned_imputed %>%
  select(-one_of(name_values))

```


Random Forest 

```{r}
rf_DP03_0120P <- randomForest(DP03_0120P ~ ., data = hi_profile_data_cleaned_imputed_dropped_features, 
                   importance = TRUE)
```

```{r}
rf_DP03_0120P
```
View columns that are the most important in predicting the outcome

```{r}
ImpData <- as.data.frame(importance(rf_DP03_0120P))
ImpData$Var.Names <- row.names(ImpData)
```


View descriptions by joining ImpData with v16 percentages

```{r}
ImpData$name <- rownames(ImpData)

ImpData_Vars <- left_join(ImpData, v16_percentages, by = "name")
```

*Need to Drop Some Variables*



Clustering!! (computing clustert output similiarity, homogeneity and completeness)




```{r}
hi_profile_data_cleaned_imputed_rownames <- hi_profile_data_cleaned_imputed %>% remove_rownames %>% 
  column_to_rownames(var="NAME")
```


*Look further into scale()*


```{r}
hi_profile_data_cleaned_imputed_rownames <- as.data.frame(scale(hi_profile_data_cleaned_imputed_rownames))
```

Determine number of clusters we put into the algorithm

```{r}
fviz_nbclust(hi_profile_data_cleaned_imputed_rownames, kmeans, method = "silhouette")
```


Cluster the newly imputed data (try density based)



```{r}
set.seed(3)
kmeans2 <- kmeans(hi_profile_data_cleaned_imputed_rownames, center = 3, nstart = 100)
```


```{r}
cluster_data <- as.data.frame(kmeans2$cluster)

cluster_data <- cluster_data %>%
  tibble::rownames_to_column(var = "NAME")

cluster_data <- cluster_data %>%
  rename(cluster = "kmeans2$cluster")
```

Map Clusters

```{r}
geometries <- get_acs(geography = "tract", 
                         variables = "DP03_0120P",
                         state = "HI", 
                         year = 2021,
                         geometry = TRUE)
```
```{r}
cluster_geo_data <- full_join(cluster_data, geometries, by = "NAME")

cluster_geo_data <- cluster_geo_data %>% 
  filter(GEOID != "15003981200")
```

```{r}
cluster_geo_sf <- st_as_sf(cluster_geo_data)

```


```{r}
ggplot(cluster_geo_sf, aes(fill = cluster)) + 
  geom_sf(color = NA) + 
  scale_fill_viridis_c() + 
  theme_void() + 
  labs(fill = "Cluster")
```

```{r}
table(cluster_geo_data$cluster)
```


Clustering feature importance

```{r}
# Access cluster centroids
centroids <- kmeans2$centers

# Create an empty data frame to store cluster means
cluster_means_df <- data.frame(Cluster = 1:3)  # Initialize with Cluster column

# For each feature, calculate the mean value within each cluster
for (col in 1:ncol(centroids)) {
  feature_name <- colnames(centroids)[col]
  cluster_means <- tapply(hi_profile_data_cleaned_imputed_rownames[[feature_name]], kmeans2$cluster, mean)
  # Add cluster means as a column to the data frame
  cluster_means_df[feature_name] <- cluster_means
}

# Now, cluster_means_df contains the mean values for each feature within each cluster


```


```{r}
cluster_means_df_longer <- cluster_means_df %>% 
  pivot_longer(!Cluster,
               names_to = "name",
               values_to = "value") 
  

# Pivot the data to have separate columns for each cluster
pivoted_data <- cluster_means_df_longer %>%
  pivot_wider(names_from = Cluster, values_from = value)

# Calculate the differences between all clusters for each feature
feature_differences <- pivoted_data %>%
  mutate(Difference_1_2 = `1` - `2`,
         Difference_1_3 = `1` - `3`,
         Difference_2_3 = `2` - `3`)


ImpCluster_Vars <- left_join(feature_differences, v16_percentages, by = "name")
```

